---
title: "Linear regression involving factors, part 1"
author: "Wing-Ho Ko"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lmtest)
library(car)
library(palmerpenguins)
```

```{r}
penguins_sim <- penguins |> select(species:sex) |> drop_na()
```

```{r}
penguins_sim
```

```{r}
pg_lm_sex <- lm(body_mass_g ~ sex, penguins_sim)
```

```{r}
summary(pg_lm_sex)
```

Checking the coding used

```{r}
pg_lm_sex$contrasts
```

Note that the first level in sex is being treated as the "control" level

```{r}
levels(penguins$sex)
```

Perform linear regression on the species variable

```{r}
pg_lm_sp <- lm(body_mass_g ~ species, penguins_sim)
```

```{r}
summary(pg_lm_sp)
```

Again note that the first level in species is being treated as the "control" level

```{r}
levels(penguins$species)
```

Use a different coding. Note that even though the problem is essentially the same, the fitted coefficients will be different

```{r}
pg_lm_sp2 <- lm(
  body_mass_g ~ species, penguins_sim, 
  contrasts=list(species=contr.sum)
)
```

```{r}
summary(pg_lm_sp2)
```

```{r}
pg_lm_sp2$contrasts
```

Functions such as `coef()`, `confint()` and `predict()` continues to work:

```{r}
confint(pg_lm_sp)
```

```{r}
pg_new <- data.frame(species=c("Adelie", "Chinstrap", "Gentoo"))
pg_pred <- predict(pg_lm_sp, pg_new)
pg_new <- cbind(pg_new, data.frame(body_mass_g = pg_pred))
```

```{r}
pg_new
```

The "prediction" is the same as the mean from each species

```{r}
penguins_sim |> group_by(species) |> summarize(
  body_mass_g = mean(body_mass_g)
)
```

Perform ANOVA using the linear regression performed

```{r}
anova(pg_lm_sp)
```

Directly perform the ANOVA using `aov()`

```{r}
pg_aov_sp <- aov(body_mass_g ~ species, penguins_sim)

summary(pg_aov_sp)
```

Test for homoscedasticity:

```{r}
leveneTest(body_mass_g ~ species, penguins_sim)
```

Test for normality

```{r}
shapiro.test(filter(penguins_sim, species=="Adelie")$body_mass_g)
```

```{r}
shapiro.test(filter(penguins_sim, species=="Chinstrap")$body_mass_g)
```

```{r}
shapiro.test(filter(penguins_sim, species=="Gentoo")$body_mass_g)
```

Perform the Kruskal-Wallis test instead of ANOVA

```{r}
kruskal.test(body_mass_g ~ species, penguins_sim)
```

Perform Tukey HSD test for pairwise comparison

```{r}
TukeyHSD(pg_aov_sp)
```

