---
title: "Linear regression involving factors, part 2"
author: "Wing-Ho Ko"
date: "2025-05-04"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(car)
library(palmerpenguins)
```

### Two-way ANOVA

Again start with the penguins dataset

```{r}
penguins_sim <- penguins |> drop_na()
```

```{r}
penguins_sim
```

Perform linear regression on the dependence of body mass on sex and species (with interactions):

```{r}
pg_lm_2cat <- lm(
  body_mass_g ~ sex * species, penguins, 
  contrasts = list(sex=contr.sum, species=contr.sum)
)

summary(pg_lm_2cat)
```

Perform type III ANOVA:

```{r}
Anova(pg_lm_2cat, type=3)
```

Note that the interaction between sex and species is significant, so type III ANOVA is the appropriate choice here

We can run the usual diagnostics for linear models:

```{r}
ggplot(penguins_sim) + geom_point(
  aes(x=species, color=sex, y=resid(pg_lm_2cat)), 
  position=position_dodge(width=0.5)
)
```

```{r}
shapiro.test(resid(pg_lm_2cat))
```

```{r}
leveneTest(pg_lm_2cat)
```

Similarly, we can obtain prediction and confidence interval as usual

```{r}
penguins_new <- distinct(penguins_sim, species, sex)
penguins_pred <- cbind(
  penguins_new, 
  predict(pg_lm_2cat, penguins_new, interval="confidence")
)
```

```{r}
penguins_pred
```

Compare with the mean of each unique combination

```{r}
penguins_sim |> group_by(sex, species) |> summarize(
  body_mass_g = mean(body_mass_g)
)
```


### Model with both cateogrical and continuous variables

Again consider penguins body mass, this time using flipper length and species as explanatory variables.

```{r}
pg_lm_2vars <- lm(
  body_mass_g ~ flipper_length_mm * species, penguins_sim, 
  contrasts=list(species=contr.sum)
)

summary(pg_lm_2vars)
```

To test hypotheses regarding species as a group, we need to use the `Anova()` function again. Similar to using sum-to-zero contrasts, we should fit with mean-subtracted flipper length

```{r}
penguins_sub <- penguins_sim |> mutate(
  flipper_diff_mm = flipper_length_mm - mean(flipper_length_mm)
)
```

```{r}
pg_lm_2sub <- lm(
  body_mass_g ~ flipper_diff_mm * species, penguins_sub,
  contrasts=list(species=contr.sum)
)
```

```{r}
summary(pg_lm_2sub)
```

```{r}
Anova(pg_lm_2sub, type=3)
```

making 3 prediction lines, one for each species

```{r}
penguins_2sub_new <- complete(
  data.frame(flipper_length_mm=NA, species=NA),
  flipper_length_mm = seq(175, 231, 2.5),
  species = distinct(penguins_sim, species)$species
) |> drop_na()
```

```{r}
penguins_2sub_new <- penguins_2sub_new |> mutate(
  flipper_diff_mm = flipper_length_mm - mean(penguins_sim$flipper_length_mm)
)
```

```{r}
penguins_2sub_pred <- cbind(
  penguins_2sub_new,
  data.frame(body_mass_g = predict(pg_lm_2sub, penguins_2sub_new))
)
```

```{r}
ggplot(mapping=aes(x=flipper_length_mm, y=body_mass_g, color=species)) +
  geom_point(aes(shape=species), data = penguins_sub) +
  geom_line(aes(linetype=species), data = penguins_2sub_pred, linewidth=1.2)
```

