---
title: "Week 7 exercises"
author: "Wing-Ho Ko"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(car)
library(lmtest)
library(statmod)
library(ciTools)
```

### Exercise 1

A record of the passengers on the Titanic is found in "Titanic.csv" in the data folder.

```{r}
titanic_raw <- read_csv("data/Titanic.csv")
```

1. Clean the data by removing any records for which any of the column has an unknown values

2. Make a box plot of the age of passengers against their fare class (PClass)

3. Perform a one-way ANOVA to check if there is statistically significant difference in the mean ages of the 3 groups

4. Perform a Tukey HSD test as a follow up. Does the Tukey HSD agree with the ANOVA?

5. Perform a Levene's test to check if ANOVA was a good idea.

6. Perform a Kruskal-Wallis test instead. Does our conclusion change as compared to ANOVA?


### Exercise 2

Return to the Titanic dataset. Again include only complete records. This time consider both sex and fare class

1. Make a box plot of age against each combination of sex and fare class (there are multiple ways to do this. Pick one that make sense to you)

2. Consider the interactive versus non-interactive model that describe how ages differ between different sex and fare class. Determine which model is better, balance between goodness-of-fit and number of parameters

3. Perform a two-way ANOVA on the non-additive model to see how age differs between different sex and fare class. Determine if interaction between sex and fare class is statistically significant.


### Exercise 3

We will look at the fish market dataset from the "fish.csv" file in the data folder.

```{r}
fish_raw <- read_csv("data/fish.csv")
```

1. Filter the dataset to include only the "species" Bream, Perch, and Pike

2. Construct a non-additive linear model where Length1 and Species are used to predict the weight of the fish, and print out its summary.

3. Use the `Anova()` function from `car` to check if the interaction between species and length is statistically significant

4. Construct a scatter plot of fish lengths (horizontal axis) versus weight (vertical axis), with shape and color of marker indicating species.

5. Use the `predict()` function to create the best-fitted line for each species of fish, and include these in the above plot.

6. Find the slope of the 3 best-fitted line you construct in part 5


### Exercise 4

Let's return to the titanic dataset. Again we'll include only complete observations

1. Perform a logistic regression on the titanic data to model how survival may depend on age, fare class, and the interaction between the two. Print out the resulting summary.

2. Perform an analysis of deviance (via `Anova()`) to check if the interaction between age and fare class is statistically significant.

3. Plot the quantile residuals of logistic regression against age, using color and shape to distinguish between different fare classes. Perform an overall (include all Age and fare class) Shapiro-Wilk test to check if the quantile residuals are normally distributed.

4. Make prediction on the odds of survival as function of both age and fare class, also include the 90% confidence interval. Plot the dependence of survival odds on age and their confidence intervals for each of the 3 classes.

5. Find the age that corresponds to 50% survival chance for both the 1st and 2nd fare class. Include 90% confidence interval (hint: check the documentation of `LC_logit()` to see how to avoid taking the logarithm)